---
title: "Calculations and charts - Italian provinces (by region) - COVID-19 Epidemic"
author: "Marcelo Eduardo Redoschi"
output:
  html_document: default
  pdf_document: default
---

```{r echo=FALSE, message=FALSE}
start_time<-Sys.time()
```

**`r format(start_time, "%a %d %b %Y %X %Z")`**

#### Based on the Protezione Civile Dataset and adjusted for the ISTAT (National Statistics Institute) regional and provincial nomenclature

*Source: dpc-covid19-ita-province.csv, own calculations*


---
```{r echo=FALSE, message=FALSE}
# Initial setup (Libraries and paths)
library(readr)
library(lubridate)
library(dplyr)
library(markdown)
library(ggplot2)
source(file.path(getwd(),"helper_functions.R"))
source(file.path(getwd(),"chart_helpers.R"))

main_dir<-file.path(getwd(), "..") # Project's main directory (top level)
notebooks_dir<-file.path(main_dir, "notebooks") 
csv_input_dir<-file.path(main_dir, "csv_input") 

pre_processing_dir<-file.path(main_dir,"optional_pre_processing") 
#raw_csv_input_dir<-file.path(pre_processing_dir, "raw_csv_input") 
prepared_data_dir<-file.path(pre_processing_dir, "prepared_data_csv") 
datasets_path<-file.path(main_dir, "Protezione-Civile-Dataset","COVID-19") 
csv_exports_dir<-file.path(main_dir, "csv_exports") 
create_directory_if_needed(csv_exports_dir)
```


```{r echo=FALSE, message=FALSE}
# ISTAT (National Statistics Institute) regional and provincial classification
ISTAT_path<-file.path(getwd(),"..", "ISTAT-CSV-Data") 
#istat_pop_file_name<-file.path(ISTAT_path,"DCIS_POPRES1_11042020115451140.csv")

istat_pop_input_file_name<-file.path(prepared_data_dir,"istat_population.csv")

if (!file.exists(istat_pop_input_file_name)) {

print(csv_input_dir)
unzip(istat_pop_zip_file_name, exdir = csv_input_dir)

list.files(csv_input_dir)
  
}


istat_area_file_name<-file.path(csv_input_dir,"DCCV_CARGEOMOR_ST_COM_12042020123304098.csv")

istat_population<-import_csv(istat_pop_input_file_name)
istat_population<-population_intl_column_names(istat_population)

istat_country<-retrieve_istat_country(istat_population)
istat_zones<-retrieve_istat_zones(istat_population)
istat_regions<-retrieve_istat_regions(istat_population)
istat_provinces<-retrieve_istat_provinces(istat_population) 
istat_provinces$regional_territory_code<-substring(istat_provinces$territory_code,1,4)
istat_provinces<-provinces_regional_territory_code_fix(istat_provinces)

istat_area<-import_csv(istat_area_file_name)
istat_area<-istat_area[istat_area$`Data.type`=="total area (km2)",]
#istat_area<-select(istat_area, ITTER107, Territory, Value)
istat_area<-select(istat_area, Territory, Value)
col_names<-colnames(istat_area)
#col_names<-gsub("ITTER107","territory_code",col_names)
col_names<-gsub("Territory","territory_name",col_names)
col_names<-gsub("Value","km2",col_names)
colnames(istat_area)<-col_names
```

```{r echo=FALSE, message=FALSE, warning=FALSE}
# Provincial data - CSV Import from local subdirectory

provincial_data_csv_fname<-file.path(datasets_path,"dati-province","dpc-covid19-ita-province.csv")
dpc_covid19_ita_provinces<-import_csv(provincial_data_csv_fname)
dpc_covid19_ita_provinces<-forli_cesena_accent_fix(dpc_covid19_ita_provinces)

provinces_data<-distinct(filter_extraneous_rows(dpc_covid19_ita_provinces))
#provinces_data<-trentino_sudtirol_fix(provinces_data,"Trentino Alto Adige / Südtirol")

provinces_data$denominazione_regione<- gsub("Friuli Venezia Giulia",'Friuli-Venezia Giulia',provinces_data$denominazione_regione)

provinces_data$denominazione_regione<- gsub("P.A. Bolzano","Trentino Alto Adige / Südtirol",provinces_data$denominazione_regione)

provinces_data$denominazione_regione<- gsub("P.A. Trento","Trentino Alto Adige / Südtirol",provinces_data$denominazione_regione)

provinces_data$denominazione_regione<- gsub("Valle d'Aosta","Valle d'Aosta / Vallée d'Aoste",provinces_data$denominazione_regione)

# setdiff(sort(unique(provinces_data$denominazione_provincia)),sort(istat_provinces$province_name))
# [1] "Aosta"                              "Bolzano"                            "Fuori Regione / Provincia Autonoma" "Massa Carrara"  
#provinces_data$denominazione_regione<-fix_region_names(istat_regions, provinces_data)

provinces_data$denominazione_provincia<- gsub("Aosta","Valle d'Aosta / Vallée d'Aoste",provinces_data$denominazione_provincia)
provinces_data$denominazione_provincia<- gsub("Bolzano","Bolzano / Bozen",provinces_data$denominazione_provincia)
provinces_data$denominazione_provincia<- gsub("Massa Carrara","Massa-Carrara",provinces_data$denominazione_provincia)

provinces_data[provinces_data$denominazione_regione=="Trentino Alto Adige / Südtirol",]$codice_regione<-4 # Fixed an inconsistency from the DPC dataset (removed the 'artificial' 21 and 22 regional codes) which caused repeated plots (impacted the "local_province_names") for the "Trentino Alto Adige / Südtirol" region.

#provinces_data<-dpc_to_istat_province_name_fix(provinces_data,istat_provinces)
```

```{r echo=FALSE, message=FALSE, warning=FALSE}
# Density
province_areas<-distinct(inner_join(istat_provinces, istat_area, by=c('province_name'='territory_name')))
province_areas$density<-province_areas$population/province_areas$km2
province_areas<-select(province_areas,"province_name","km2","density") 
provinces_data<-inner_join(provinces_data, province_areas,by=c('denominazione_provincia'='province_name'))
```

```{r echo=FALSE, message=FALSE, warning=FALSE}
provinces_regions_key <-
  inner_join(istat_provinces,
  provinces_data,
  by = c('province_name' = 'denominazione_provincia'))
  
  provinces_regions_key <-
  distinct(
  select(
  provinces_regions_key,
  province_name,
  sigla_provincia,
  codice_provincia,
  territory_code,
  denominazione_regione,
  codice_regione,
  regional_territory_code
  )
  )
  italian_regions <-
  distinct(select(provinces_regions_key, denominazione_regione))
  italian_regions <- arrange(italian_regions, denominazione_regione)
```

```{r echo=FALSE, message=FALSE, warning=FALSE}
provinces_data <-
  inner_join(provinces_data,
             istat_provinces,
             by = c('denominazione_provincia' = 'province_name'))
             provinces_data <- add_dt_column(provinces_data)
```


```{r echo=FALSE, message=FALSE}
# Adds total case columns for the respective dates (generalized version)
#provinces_data <-
#  expand_dt_columns(provinces_data, "denominazione_provincia", "totale_casi") 
```

```{r echo=FALSE, message=FALSE}
provinces_calculations <- provinces_data

provinces_calculations <-provinces_calculations[provinces_calculations$dt>'2021-1-1',]

current_total_cases <-
total_cases_on_dt(provinces_data, days_before(provinces_data, 0))
total_cases_1_day_before <-
total_cases_on_dt(provinces_data, days_before(provinces_data, 1))
total_cases_2_days_before <-
total_cases_on_dt(provinces_data, days_before(provinces_data, 2))
total_cases_3_days_before <-
total_cases_on_dt(provinces_data, days_before(provinces_data, 3))
total_cases_4_days_before <-
total_cases_on_dt(provinces_data, days_before(provinces_data, 4))

provinces_calculations$current_rate <-
current_total_cases / total_cases_1_day_before
provinces_calculations$yesterdays_rate <-
total_cases_1_day_before / total_cases_2_days_before
provinces_calculations$rate_two_days_before <-
total_cases_2_days_before / total_cases_3_days_before
provinces_calculations$rate_three_days_before <-
total_cases_3_days_before / total_cases_4_days_before

provinces_calculations$cubic_rate <-
'^'(current_total_cases / total_cases_3_days_before,1/3)
provinces_calculations$cubic_perc_rate <-
(provinces_calculations$cubic_rate - 1) * 100
provinces_calculations$current_perc_rate <-
(provinces_calculations$current_rate - 1) * 100
provinces_calculations$trend <-
(provinces_calculations$current_rate - provinces_calculations$cubic_rate)*100
provinces_calculations$per_capita_rate <-
provinces_calculations$totale_casi / provinces_calculations$population *
1000

most_recent_dt <- retrieve_most_recent_dt(provinces_calculations)
three_days_before <- most_recent_dt - 3

provinces_trends <-
provinces_calculations[provinces_calculations$dt == most_recent_dt, ]

provinces_calculations<-distinct(provinces_calculations)

provinces_trends<-distinct(select(
provinces_trends,
denominazione_provincia,
totale_casi,
current_perc_rate,
trend,
per_capita_rate
))

provinces_trends<-translated_province_column_names(provinces_trends)

provinces_trends<-revised_province_column_names(provinces_trends)
provinces_trends<-rounded(provinces_trends,c('daily_growth','per_capita','trend'),2)

#most_recent_dt<-max(provinces_calculations$dt)
```

### Province trends observed on `r most_recent_dt`
```{r echo=FALSE, message=FALSE}
province_stats<-summary(select(provinces_trends,daily_growth, per_capita, trend, total_cases))
#print(most_recent_dt, row.names=F)
print(province_stats)
```

#### With an augmenting rate of increase in the number of cases
```{r echo=FALSE, message=FALSE}
prov_increasing_trend <- provinces_trends[provinces_trends$trend > 0, ] # increasing
prov_increasing_trend <- arrange(prov_increasing_trend,-trend)

print(prov_increasing_trend)
prov_increasing_trend_sel<-select(prov_increasing_trend,daily_growth, per_capita, trend, total_cases)
print(summary(prov_increasing_trend_sel))
```

#### With a steady or declining rate of increase 
```{r echo=FALSE, message=FALSE} 
prov_declining_trend <- provinces_trends[provinces_trends$trend <= 0, ]
prov_declining_trend <- arrange(prov_declining_trend,-trend)

print(prov_declining_trend)
prov_declining_trend_sel<-select(prov_declining_trend,daily_growth, per_capita, trend, total_cases)
print(summary(prov_declining_trend_sel))
```
*Trend is an approximation = (current_rate-cubic_rate)x100*


```{r echo=FALSE, message=FALSE} 
### Total cases per area (km2)
```

### Provinces per capita (expressed as per thousand people)  ‰ rate.
```{r echo=FALSE, message=FALSE} 
provinces_per_capita <-
  select(
  provinces_calculations,
  dt,
  denominazione_provincia,
#  denominazione_regione,
  totale_casi,
  per_capita_rate
  )
  most_recent_dt <- max(provinces_per_capita$dt)
  provinces_per_capita <- arrange(provinces_per_capita, -per_capita_rate)
  #print(most_recent_dt)
  
  provinces_per_capita_latest <-
  provinces_per_capita[provinces_per_capita$dt == most_recent_dt, ]
  provinces_per_capita_latest$pos <-
  nrow(provinces_per_capita_latest) - rank(provinces_per_capita_latest$per_capita_rate) +
  1
  prov_per_capita_latest_sel <-
  arrange(provinces_per_capita_latest,-per_capita_rate)
  provinces_per_capita_latest_sel<-select(
  provinces_per_capita_latest,
  pos,
  denominazione_provincia,
#  denominazione_regione,
  per_capita_rate
  )
  
#  prov_per_capita_latest_sel$per_capita_rate<-round(prov_per_capita_latest_sel$per_capita_rate, digits=2)
  
  prov_per_capita_latest_sel<-rounded(prov_per_capita_latest_sel, c("per_capita_rate"), 2)
    
  prov_per_capita_latest_sel$dt<-NULL 

  prov_per_capita_latest_sel<-translated_province_column_names(prov_per_capita_latest_sel)    
  prov_per_capita_latest_sel<-revised_province_column_names(prov_per_capita_latest_sel)

  print(prov_per_capita_latest_sel, row.names = F)
    
  rm(prov_per_capita_latest_sel)
```

### Provinces - total cases per density (population per square kilometer).
```{r echo=FALSE, message=FALSE} 

provinces_calculations$total_case_density<-provinces_calculations$totale_casi/provinces_calculations$density
provinces_per_density <-
  select(
  provinces_calculations,
  dt,
  denominazione_provincia,
  totale_casi,
  density,
  total_case_density
  )
  most_recent_dt <- max(provinces_per_density$dt)
  provinces_per_density <- arrange(provinces_per_density, -total_case_density)
  #print(most_recent_dt)
  
  provinces_per_density_latest <-
  provinces_per_density[provinces_per_density$dt == most_recent_dt, ]
  provinces_per_density_latest$pos <-
  nrow(provinces_per_density_latest) - rank(provinces_per_density_latest$total_case_density) +
  1
  prov_per_density_latest_sel <-
  arrange(provinces_per_density_latest,-total_case_density)
  provinces_per_density_latest_sel<-select(
  provinces_per_density_latest,
  pos,
  denominazione_provincia,
#  denominazione_regione,
  total_case_density
  )
  
#  prov_per_density_latest_sel$total_case_density<-round(prov_per_density_latest_sel$total_case_density, digits=2)
  
  prov_per_density_latest_sel<-rounded(prov_per_density_latest_sel, c("total_case_density"), 2)
    
  prov_per_density_latest_sel$dt<-NULL 

  prov_per_density_latest_sel<-translated_province_column_names(prov_per_density_latest_sel)    
  prov_per_density_latest_sel<-revised_province_column_names(prov_per_density_latest_sel)

  print(prov_per_density_latest_sel, row.names = F)
    
  rm(prov_per_density_latest_sel)
```

```{r echo=FALSE, message=FALSE} 
write.csv(provinces_calculations,file.path(csv_exports_dir,"provinces_calculations.csv"))
write.csv(provinces_trends,file.path(csv_exports_dir,"provinces_trends.csv"))
write.csv(prov_increasing_trend,file.path(csv_exports_dir,"provinces_increasing_trend.csv"))
write.csv(prov_declining_trend,file.path(csv_exports_dir,"provinces_declining_trend.csv"))
write.csv(provinces_regions_key,file.path(csv_exports_dir,"provinces_regions_key.csv"))
write.csv(provinces_per_capita,file.path(csv_exports_dir,"provinces_per_capita.csv"))
write.csv(provinces_per_capita_latest,file.path(csv_exports_dir,"provinces_per_capita_latest.csv"))
```

### Graphs - Total case progression - Provinces by region (for the most recent days observed)

```{r echo=FALSE, message=FALSE}
total_cases_by_province<-select(provinces_data, dt, denominazione_provincia, denominazione_regione, totale_casi)
total_cases_by_province<-translated_province_column_names(total_cases_by_province)

# sorted by region name
total_cases_by_province<-arrange(total_cases_by_province, denominazione_regione, province_name, total_cases)

```

```{r  fig.height=5, fig.width=10, echo=FALSE, message=FALSE}
require(ggplot2)

specified_cols <-
c(
"dt",
"denominazione_provincia",
"denominazione_regione",
"totale_casi",
"current_perc_rate",
"cubic_perc_rate",
"per_capita_rate"
)

prov_calc_sel <-
select(provinces_calculations, all_of(specified_cols))

prov_calc_sel <- filtered_for_the_most_recent_days(prov_calc_sel, 1)

prov_calc_sel <- translated_province_column_names(prov_calc_sel)

# restricted by date
provinces_calculations<-provinces_calculations[provinces_calculations$dt>'2020-4-15',]

```

```{r echo=FALSE, message=FALSE}
provinces_calculations_most_recent_dt<-provinces_calculations[provinces_calculations$dt==most_recent_dt,]
histogram_title<-paste("Provinces histogram - growth in new cases",most_recent_dt)
ggplot(provinces_calculations_most_recent_dt, aes(cubic_perc_rate))+geom_histogram(color="yellow",fill="#5a80b2")+labs(title=histogram_title, x="Weighted average percent rate for previous three days",y = "Number of provinces")
```

```{r  fig.height=5, fig.width=10, echo=FALSE, message=FALSE}

for (i in 1:nrow(italian_regions)) {

selected_region <- italian_regions[i,]

selected_region_name <- as.character(selected_region)

local_provinces_key <-
provinces_regions_key[provinces_regions_key$denominazione_regione == selected_region_name, ]
local_province_names <- local_provinces_key$province_name

# Local provinces 

  local_province_data <- provinces_calculations[provinces_calculations$denominazione_regione == selected_region, ]
  local_province_data <-
    arrange(local_province_data, denominazione_provincia)

# Regional chart (one or more provinces)

current_rate_color<-"#EB9B5E"
cubic_rate_color<-"#2C9FD4"
x_label_text="Province"
cubic_rate_line_type="dashed"

regional_provinces_cubic_rate_chart <-cubic_rate_boxplot_chart(local_province_data, selected_region,current_rate_color,cubic_rate_color,x_label_text, cubic_rate_line_type)

font_size<-11
angle<-20 
regional_provinces_cubic_rate_chart<-styled_x_axis(regional_provinces_cubic_rate_chart,font_size,angle) 

### ************ Regional provinces chart ***********   
print(regional_provinces_cubic_rate_chart)
### *************************************************

local_provinces <-
filtered_by_region(prov_calc_sel, selected_region)

observation_dt <- mean(local_provinces$dt)
observation_dt_txt<-format(observation_dt, format="%d %B %Y")
regional_summary_txt<-paste("Regional summary for '",selected_region,"'")

# `r regional_summary_txt`

# `r observation_dt_txt`

print(observation_dt, row.names = F)

local_provinces$denominazione_regione <- NULL

formatted_local_province_rates<-select(local_provinces, -dt)
formatted_local_province_rates<-revised_province_column_names(formatted_local_province_rates)
formatted_local_province_rates_p<-rounded(formatted_local_province_rates,c("daily_growth","per_capita","three_day_rate"),2) # for printing
print(formatted_local_province_rates_p, row.names = F)  

# Summary statistics
local_province_rates_sel<-select(formatted_local_province_rates,daily_growth,three_day_rate,total_cases,per_capita)

if (nrow(formatted_local_province_rates)>1) {

  print(summary(local_province_rates_sel))

}

# Provincial charts (current vs. three and seven day averages)

for (j in 1:length(local_province_names)) {

local_province_name <- local_province_names[j]

local_province_data <-
total_cases_by_province[total_cases_by_province$province_name == local_province_name,]

local_province_data<-add_three_and_seven_day_rate_columns(local_province_data,"total_cases")

local_province_rates <-
local_province_data[is.finite(local_province_data$seventh_perc_rate),]

combined_chart<-provincial_combined_rates_chart(local_province_rates, local_province_name)

num_days_in_the_past<-30 
starting_dt<-max(local_province_rates$dt)-num_days_in_the_past 
combined_chart_recent<-provincial_combined_rates_from_dt_chart(local_province_rates, local_province_name, starting_dt)


print(combined_chart)
print(paste("Shown in more detail: recent observations, for the past",num_days_in_the_past,"days"), row.names = FALSE, quote=FALSE)
print(combined_chart_recent)
rm(combined_chart)

rm(combined_chart_recent)
}

rm(local_province_rates)
rm(local_provinces)

} # regions loop

```

```{r echo=FALSE, message=FALSE}
finish_time<-Sys.time()
```

---
```{r echo=FALSE, message=FALSE}
finish_time<-Sys.time()
running_time<-as.POSIXct(finish_time)-as.POSIXct(start_time)

```
*This document's processing time: `r running_time`*
